Estimating the dynamics of SPARQL query results\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}

### 2019-04-11 kgswc19.tex
second

\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
    colorlinks, linkcolor={dark-red},
    citecolor={dark-green}, urlcolor={dark-blue},
    pdftitle={Estimating the dynamics of SPARQL query results over Wikidata},    % title
    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={KGSWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of\\SPARQL query results over Wikidata\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IMFD; DCC, University of Chile\\
\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We propose a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates.  \ah{In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.}

@@

\begin{table*}[t]
	\begin{tabular}{lrrrrrrr}    \toprule
		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
		\textit{Random Baseline}   & 0.5002262 & 0.5126697 & 0.5027149 & 0.511086  & 0.5031674 & 0.5063348 & 0.4979638 \\
		Decision Trees    & 0.4708145 & 0.5233032 & 0.7821267 & 0.4857466 & 0.5671946 & 0.6803167 & 0.6047511 \\
		Naive Bayes       & 0.5169683 & 0.4647059 & 0.7638009 & 0.5242081 & 0.5420814 & 0.7588235 & 0.5337104 \\
		Nearest Neighbors & 0.4846154 & 0.5106335 & 0.7819005 & 0.5056561 & 0.6776018 & 0.7760181 & 0.6832579 \\
		Linear SVM        & 0.5914027 & 0.5108597 & 0.7638009 & 0.5914027 & 0.7751131 & 0.7647059 & 0.7746606 \\    \midrule
		\textbf{Better}            & 0.5914027 & 0.5233032 & 0.7821267 & 0.5914027 & 0.7751131 & 0.7760181 & 0.7746606 \\    \bottomrule
		\textit{Random Baseline}   & 0.5010055 & 0.5005028 & 0.4959779 & 0.4987431 & 0.4974862 & 0.4856712 & 0.4816491 \\
		Decision Trees    & 0.4497235 & 0.5294118 & 0.782554  & 0.4954751 & 0.5723982 & 0.688537  & 0.6289593 \\
		Naive Bayes       & 0.5160885 & 0.4705882 & 0.7946204 & 0.5160885 & 0.5510307 & 0.7787833 & 0.5334339 \\
		Nearest Neighbors & 0.5125691 & 0.5296631 & 0.7868276 & 0.5115636 & 0.6792358 & 0.7875817 & 0.6726998 \\
		Linear SVM        & 0.5957768 & 0.5075415 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \midrule
		\textbf{Better}            & 0.5957768 & 0.5296631 & 0.7946204 & 0.5849673 & 0.7895928 & 0.7885872 & 0.7880845 \\    \bottomrule
		\textit{Random Baseline}   & 0.4961215 & 0.5087266 & 0.5054945 & 0.487395  & 0.512605  & 0.4938591 & 0.5090498 \\
		Decision Trees    & 0.448287  & 0.54234   & 0.7918552 & 0.4974144 & 0.6263736 & 0.7117001 & 0.6764706 \\
		Naive Bayes       & 0.5258565 & 0.4605688 & 0.7983193 & 0.5235941 & 0.5688429 & 0.7685844 & 0.5585003 \\
		Nearest Neighbors & 0.5206852 & 0.5336134 & 0.789916  & 0.5310278 & 0.6706529 & 0.7886231 & 0.6780866 \\
		Linear SVM        & 0.5727214 & 0.5       & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \midrule
		\textbf{Better}          & 0.5727214 & 0.54234   & 0.7983193 & 0.5688429 & 0.7957337 & 0.7950873 & 0.7947641 \\    \bottomrule
		\textit{Random Baseline}   & 0.4766214 & 0.4962293 & 0.5067873 & 0.5030166 & 0.4788839 & 0.484917  & 0.5030166 \\
		Decision Trees    & 0.4969834 & 0.4819005 & 0.8205128 & 0.5346908 & 0.7292609 & 0.7503771 & 0.7277526 \\
		Naive Bayes       & 0.5090498 & 0.4834087 & 0.8280543 & 0.5037707 & 0.5791855 & 0.7978884 & 0.5874811 \\
		Nearest Neighbors & 0.5580694 & 0.4871795 & 0.8288084 & 0.5671192 & 0.739819  & 0.8220211 & 0.7239819 \\
		Linear SVM        & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\    \midrule
		\textbf{Better}            & 0.5656109 & 0.5098039 & 0.8288084 & 0.5791855 & 0.8273002 & 0.8288084 & 0.8295626 \\     \bottomrule
	\end{tabular}
\end{table*}

>>>>

\usepackage{xcolor}
\usepackage{hyperref}
\definecolor{dark-blue}{rgb}{0.0,0.0,0.2}
\definecolor{dark-green}{rgb}{0.0,0.2,0.0}
\definecolor{dark-red}{rgb}{0.2,0.0,0.0}
\hypersetup{
    colorlinks, linkcolor={dark-red},
    citecolor={dark-green}, urlcolor={dark-blue},
    pdftitle={Estimating the dynamics of SPARQL query results},    % title
    pdfauthor={Alberto Moya, Aidan Hogan},     % author
    pdfsubject={KGSWC 2019},   % subject of the document
    pdfkeywords={SPARQL;} {Linked Data;} {Dynamics;} {Wikidata;}, % list of keywords
}

@@

\begin{document}
%
\title{Estimating the dynamics of SPARQL query results\thanks{This work was supported by CONICYT PFCHA, Doctorado Becas Chile/2017 -- 72180000, the Millennium Institute for Foundational Research on Data (IMFD) and FONDECYT Grant No. 1181896.}}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Alberto Moya\orcidID{0000-0002-7003-5087} \and 
Aidan Hogan\orcidID{0000-0001-9482-1982}}
%
\authorrunning{A. Moya and A. Hogan}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{IMFD; DCC, University of Chile\\
\email{\{amoya,ahogan\}@dcc.uhile.cl}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
We address the problem of estimating when the results of an input SPARQL query results over dynamic RDF datasets will change. We propose a framework that extracts features from the query and/or from past versions of the target dataset. The resulting features are then used to predict (1) whether or not the results for a query will change in the near future using classifiers, and (2) when they might change (if at all) using linear regression. To evaluate our proposal, we create a gold standard based on 23 versions of Wikidata and a curated collection of 221 SPARQL queries. Our results show that the quality of predictions using query features and lightweight statistics of the dynamics of individual predicates is competitive with that obtained using (more costly to derive) knowledge of the complete historical changes in the query results.

@@

In the case of predicate dynamics, as discussed in Section~\ref{sec:approach}, there are potentially many predicates per query, each with its own value for $\Delta(\mathcal{G},p)$, where to reduce (and fix) dimensionality, we may apply an aggregate function to choose the min, max or mean value over all predicates. In preliminary experiments, we found the mean value to offer the best results, followed by the max value; hence in what follows we adopt the mean value in our experiments.

@@

\begin{table*}[t]
	\begin{tabular}{lrrrrrrr}    \toprule
		\textbf{Classifier}        & \textsc{Q}         & \textsc{P}         & \textsc{R}         & \textsc{QP}        & \textsc{QR}        & \textsc{PR}        & \textsc{QPR}       \\    \midrule
		\textit{Random Baseline}   & 0.490  & 0.486 & 0.500   & 0.502 & 0.486 & 0.495 & 0.491 \\
		Decision Trees    & 0.471 & 0.513 & \underline{0.782} & 0.497 & 0.572 & 0.705 & 0.605 \\
		Naive Bayes       & 0.514 & \underline{0.523} & 0.764 & 0.514 & 0.540 & 0.763 & 0.540  \\
		Nearest Neighbors & 0.505 & 0.521 & 0.778 & 0.509 & 0.679 & \underline{0.781} & 0.680  \\
		Linear SVM        & \underline{0.596} & 0.518 & 0.764 & \underline{0.593} & \underline{0.775} & 0.767 & \underline{0.774} \\    \midrule
		\textbf{Better}   & 0.596 & 0.523 & 0.782 & 0.593 & 0.775 & 0.781 & 0.774 \\    \bottomrule
		\textit{Random Baseline}   & 0.487 & 0.497 & 0.500 & 0.512 & 0.502 & 0.515 & 0.501 \\
		Decision Trees    & 0.482 & 0.515 & 0.782 & 0.494 & 0.580  & 0.714 & 0.606 \\
		Naive Bayes       & 0.514 & 0.514 & 0.782 & 0.513 & 0.561 & \underline{0.797} & 0.554 \\
		Nearest Neighbors & 0.509 & \underline{0.535} & \underline{0.797} & 0.525 & 0.676 & 0.791 & 0.674 \\
		Linear SVM        & \underline{0.600} & 0.516 & 0.785 & \underline{0.590} & \underline{0.788} & 0.790  & \underline{0.788} \\    \midrule
		\textbf{Better}   & 0.600 & 0.535 & 0.797 & 0.590 & 0.788 & 0.797 & 0.788 \\    \bottomrule
		\textit{Random Baseline}   & 0.507 & 0.498 & 0.487 & 0.506 & 0.486 & 0.497 & 0.495 \\
		Decision Trees    & 0.469 & 0.514 & 0.794 & 0.457 & 0.656 & 0.731 & 0.642 \\
		Naive Bayes       & 0.517 & 0.519 & \underline{0.797} & 0.516 & 0.577 & 0.791 & 0.576 \\
		Nearest Neighbors & 0.514 & \underline{0.529} & 0.787 & 0.518 & 0.654 & \underline{0.800} & 0.65  \\
		Linear SVM        & \underline{0.578} & 0.512 & \underline{0.797} & \underline{0.571} & \underline{0.796} & \underline{0.800} & \underline{0.796} \\    \midrule
		\textbf{Better}   & 0.578 & 0.529 & 0.797 & 0.571 & 0.796 & 0.800 & 0.796 \\    \bottomrule
		\textit{Random Baseline}   & 0.514 & 0.515 & 0.502 & 0.493 & 0.499 & 0.499 & 0.502 \\
		Decision Trees    & 0.512 & 0.517 & \underline{0.827} & 0.493 & 0.725 & 0.716 & 0.704 \\
		Naive Bayes       & 0.515 & \underline{0.519} & \underline{0.827} & 0.508 & 0.599 & 0.810  & 0.594 \\
		Nearest Neighbors & 0.549 & 0.481 & \underline{0.827} & 0.555 & 0.744 & 0.821 & 0.732 \\
		Linear SVM        & \underline{0.554} & \underline{0.519} & \underline{0.827} & \underline{0.560} & \underline{0.819} & \underline{0.831} & \underline{0.815} \\    \midrule
		\textbf{Better}   & 0.554 & 0.519 & 0.827 & 0.560 & 0.819 & 0.831 & 0.815 \\    \bottomrule
	\end{tabular}
\end{table*}

\begin{figure}[t]
	%\begin{minipage}[b]{0.30\linewidth}
	\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={F1},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,4},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
	};

		\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
	};

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Precision},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,23},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.49)(2,0.600)(3,0.578)(4,0.554)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.486)(2,0.535)(3,0.529)(4,0.519)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.5)(2,0.797)(3,0.797)(4,0.827)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.502)(2,0.590)(3,0.571)(4,0.560)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.486)(2,0.788)(3,0.796)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.495)(2,0.797)(3,0.800)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.491)(2,0.788)(3,0.796)(4,0.815)
	};

	\end{axis}
	\end{tikzpicture}
	%		\caption{Evolution of the number of triples on Wikidata.}
	%		\label{graph:datasize}
	%	\end{minipage}
	\hspace{0.1cm}
	%	\begin{minipage}[b]{0.30\linewidth}
	%		\centering
	\begin{tikzpicture}[thick,font=\scriptsize]
	\begin{axis}[
	%title={Temperature dependence of CuSO$_4\cdot$5H$_2$O solubility},
	xlabel={Recall},
	width=5cm, 
	height=5cm,
	ymin=0,
	xtick={1,2,...,23},
	xticklabels={},
	extra x ticks={1,23},
	legend pos=north west,
	ymajorgrids=true,
	grid style=dashed,
	]

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.596)(2,0.600)(3,0.578)(4,0.554)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.523)(2,0.535)(3,0.529)(4,0.519)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.782)(2,0.797)(3,0.797)(4,0.827)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.593)(2,0.590)(3,0.571)(4,0.560)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.775)(2,0.788)(3,0.796)(4,0.819)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.781)(2,0.797)(3,0.800)(4,0.831)
	};

	\addplot[
	color=blue,
	mark=diamond,
	]
	coordinates {
		(1,0.774)(2,0.788)(3,0.796)(4,0.815)
	};

	\end{axis}
	\end{tikzpicture}
	\caption{Evolution of total triples in Wikidata (left) and number of triples added to and removed from the $n$\textsuperscript{th} version for the $n+1$\textsuperscript{th} version (right) \label{fig:result}}
\end{figure}
